quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-08-04 paper-quollr.Rmd
Updated the paper with new version

hex_grid_with_counts_scurve <- left_join(
  hex_grid_scurve, counts_df_scurve, 
  by = c("hex_poly_id" = "h"))

hex_grid_nonempty_scurve <- hex_grid_scurve |>
  filter(hex_poly_id %in% df_bin_centroids_scurve$h)

@@

hex_grid_poly_scurve <- ggplot(
  data = hex_grid_with_counts_scurve, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = scurve_umap_scaled, 
             aes(x = emb1, y = emb2), 
             alpha = 0.2, size = 0.5, color = clr_choice) + 
  interior_annotation("b") +
  xlim(sc_xlims) +
  ylim(sc_ylims)

@@

hex_centroids_scurve <- ggplot(
  data = hex_grid_with_counts_scurve, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = hex_poly_id), 
               fill = "#ffffff")  +
  geom_point(data = df_bin_centroids_scurve, 
             aes(x = c_x, y = c_y), 
             size = 1, color = "#33a02c") +
  interior_annotation("c") +
  xlim(sc_xlims) +
  ylim(sc_ylims)

@@

```r
devtools::install_github("JayaniLakshika/quollr")
```

@@

hex_grid_temp45 <- hex_grid_temp |> 
  filter(hex_poly_id == 45)

@@

                            group = hex_poly_id),

@@

For each centroid, six vertices are computed and assigned a polygon ID (`hex_poly_id`) corresponding to the centroid’s `h`. These points are aggregated into a tibble that includes the polygon ID (`hex_poly_id`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.

@@

```{r, echo=TRUE}
find_low_dens_hex(
  centroids_data = df_bin_centroids, 
  b1 = 15, 
  benchmark_mean_dens = 0.05
)

@@

```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="jkjio"}
nldrdt_link <- show_link_plots(
  point_data = df_exe, 
  edge_data = trimesh, 
  point_colour = clr_choice
  )

@@

# hex_grid <- tsne_limb_obj$hb_obj$hex_poly
# counts_df <- tsne_limb_obj$hb_obj$std_cts
# 
# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("hex_poly_id" = "h"))
# 
# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = n_h)) +
#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = h)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed() +
#   theme_minimal()
```

>>>>

hex_grid_with_counts_scurve <- left_join(
  hex_grid_scurve, counts_df_scurve, 
  by = c("h" = "h"))

hex_grid_nonempty_scurve <- hex_grid_scurve |>
  filter(h %in% df_bin_centroids_scurve$h)

@@

hex_grid_poly_scurve <- ggplot(
  data = hex_grid_with_counts_scurve, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = h), 
               fill = "#ffffff") +
  geom_point(data = scurve_umap_scaled, 
             aes(x = emb1, y = emb2), 
             alpha = 0.2, size = 0.5, color = clr_choice) + 
  interior_annotation("b") +
  xlim(sc_xlims) +
  ylim(sc_ylims)

@@

hex_centroids_scurve <- ggplot(
  data = hex_grid_with_counts_scurve, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = h), 
               fill = "#ffffff")  +
  geom_point(data = df_bin_centroids_scurve, 
             aes(x = c_x, y = c_y), 
             size = 1, color = "#33a02c") +
  interior_annotation("c") +
  xlim(sc_xlims) +
  ylim(sc_ylims)

@@

```r
pak::pak("JayaniLakshika/quollr")
```

@@

hex_grid_temp45 <- hex_grid_temp |> 
  filter(h == 45)

@@

                            group = h),

@@

For each centroid, six vertices are computed and assigned a polygon ID (`h`) corresponding to the centroid’s `h`. These points are aggregated into a tibble that includes the polygon ID (`h`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.

@@

```{r, echo=TRUE}
find_low_dens_hex(
  model_2d = df_bin_centroids, 
  b1 = 15, 
  benchmark_mean_dens = 0.05
)

@@

```{r linkerror, echo=TRUE, eval=knitr::is_html_output(), fig.cap="Link plot showing the relationship between the NLDR layout (left) and the fitted model overlaid with the data in $7\\text{-}D$ (right)."}

nldrdt_link <- show_link_plots(
  point_data = df_exe, 
  edge_data = trimesh, 
  point_colour = clr_choice
  )

@@

# hex_grid <- tsne_limb_obj$hb_obj$hex_poly
# counts_df <- tsne_limb_obj$hb_obj$std_cts
# 
# hex_grid_with_counts <- left_join(hex_grid, counts_df, by = c("h" = "h"))
# 
# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = n_h)) +
#   #geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = h)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed() +
#   theme_minimal()
```

### 2025-08-04 paper-quollr.tex
Updated the paper with new version

\begin{verbatim}
devtools::install_github("JayaniLakshika/quollr")
\end{verbatim}

@@

\begin{verbatim}
> $scaled_nldr
> # A tibble: 5,000 x 3
>     emb1  emb2    ID
>    <dbl> <dbl> <int>
>  1 0.707 0.839     1
>  2 0.231 0.401     2
>  3 0.232 0.215     3
>  4 0.790 0.564     4
>  5 0.761 0.551     5
>  6 0.445 0.721     6
>  7 0.900 0.137     7
>  8 0.247 0.392     8
>  9 0.325 0.542     9
> 10 0.278 0.231    10
> # i 4,990 more rows
> 
> $lim1
> [1] -14.42166  13.32655
> 
> $lim2
> [1] -12.43687  12.32455
\end{verbatim}

@@

\begin{verbatim}
> $b2
> [1] 16
> 
> $a1
> [1] 0.08326136
> 
> $a2
> [1] 0.07210645
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 240 x 3
>        h     c_x     c_y
>    <int>   <dbl>   <dbl>
>  1     1 -0.1    -0.0892
>  2     2 -0.0167 -0.0892
>  3     3  0.0665 -0.0892
>  4     4  0.150  -0.0892
>  5     5  0.233  -0.0892
>  6     6  0.316  -0.0892
>  7     7  0.400  -0.0892
>  8     8  0.483  -0.0892
>  9     9  0.566  -0.0892
> 10    10  0.649  -0.0892
> # i 230 more rows
\end{verbatim}

@@

For each centroid, six vertices are computed and assigned a polygon ID (\texttt{hex\_poly\_id}) corresponding to the centroid's \texttt{h}. These points are aggregated into a tibble that includes the polygon ID (\texttt{hex\_poly\_id}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.

@@

\begin{verbatim}
>   hex_poly_id           x           y
> 1           1 -0.10000000 -0.04116510
> 2           1 -0.14163068 -0.06520059
> 3           1 -0.14163068 -0.11327155
> 4           1 -0.10000000 -0.13730704
> 5           1 -0.05836932 -0.11327155
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 4
>    emb1  emb2    ID     h
>   <dbl> <dbl> <int> <int>
> 1 0.707 0.839     1   205
> 2 0.231 0.401     2   109
> 3 0.232 0.215     3    65
> 4 0.790 0.564     4   146
> 5 0.761 0.551     5   146
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 3
>       h   n_h    w_h
>   <int> <int>  <dbl>
> 1    21    12 0.0024
> 2    22    17 0.0034
> 3    23    22 0.0044
> 4    24    10 0.002 
> 5    25     7 0.0014
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 2
>       h pts_list  
>   <int> <list>    
> 1    21 <int [12]>
> 2    22 <int [17]>
> 3    23 <int [22]>
> 4    24 <int [10]>
> 5    25 <int [7]>
\end{verbatim}

@@

\begin{verbatim}
>   h         c_x         c_y n_h w_h
> 1 1 -0.10000000 -0.08923607   0   0
> 2 2 -0.01673864 -0.08923607   0   0
> 3 3  0.06652271 -0.08923607   0   0
> 4 4  0.14978407 -0.08923607   0   0
> 5 5  0.23304543 -0.08923607   0   0
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 8
>    from    to  x_from  y_from    x_to    y_to from_count to_count
>   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
> 1     1     2 -0.1    -0.0892 -0.0167 -0.0892          0        0
> 2    16    17 -0.0584 -0.0171  0.0249 -0.0171          0        0
> 3    16    32 -0.0584 -0.0171 -0.0167  0.0550          0        0
> 4     3    17  0.0665 -0.0892  0.0249 -0.0171          0        0
> 5    17    18  0.0249 -0.0171  0.108  -0.0171          0        0
\end{verbatim}

@@

\begin{verbatim}
find_low_dens_hex(
  centroids_data = df_bin_centroids, 
  b1 = 15, 
  benchmark_mean_dens = 0.05
)
\end{verbatim}

\begin{verbatim}
>   [1]   1   2   3   4   5   6   7   8   9  10  11  12  13  14  15  16  17  18
>  [19]  19  20  21  22  23  24  25  26  27  28  29  30  31  32  33  34  35  36
>  [37]  37  38  39  40  41  42  43  44  45  46  47  48  49  50  51  52  53  54
>  [55]  55  56  57  58  59  60  61  62  63  64  65  66  67  68  69  70  71  72
>  [73]  73  74  75  76  77  78  79  80  81  82  83  84  85  86  87  88  89  90
>  [91]  91  92  93  94  95  96  97  98  99 100 101 102 103 104 105 106 107 108
> [109] 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126
> [127] 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144
> [145] 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162
> [163] 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180
> [181] 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198
> [199] 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215 216
> [217] 217 218 219 220 221 222 223 224 225 226 227 228 229 230 231 232 233 234
> [235] 235 236 237 238 239 240
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 8
>       h      x1    x2    x3         x4        x5       x6        x7
>   <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
> 1    21 -0.992   1.91  1.11 -0.000427   0.000624  0.00749  0.00105 
> 2    22 -0.906   1.93  1.41 -0.0000183  0.00331  -0.0204  -0.000363
> 3    23 -0.680   1.93  1.72 -0.000810  -0.00259  -0.00449  0.00153 
> 4    24 -0.272   1.93  1.96  0.00251    0.00668  -0.0460   0.00128 
> 5    25  0.0760  1.93  2.00  0.00876    0.00447   0.00851 -0.00195
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 4
>   pred_emb_1 pred_emb_2    ID pred_hb_id
>        <dbl>      <dbl> <int>      <int>
> 1      0.691      0.848     1        205
> 2      0.191      0.416     2        109
> 3      0.316      0.199     3         66
> 4      0.774      0.560     4        146
> 5      0.774      0.560     5        146
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 1 x 2
>   Error  RMSE
>   <dbl> <dbl>
> 1 1481. 0.180
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 32
>      ID      x1     x2       x3       x4       x5      x6        x7 pred_hb_id
>   <int>   <dbl>  <dbl>    <dbl>    <dbl>    <dbl>   <dbl>     <dbl>      <int>
> 1     1 -0.120  1.64   -1.99     0.0104   0.0125   0.0923 -0.00128         205
> 2     2 -0.0492 1.51    0.00121 -0.0177   0.00726 -0.0362 -0.00535         109
> 3     3 -0.774  1.30    0.367   -0.00173  0.0156  -0.0962  0.00335          66
> 4     4 -0.606  0.246  -1.80    -0.00897 -0.0187  -0.0716  0.00126         146
> 5     5 -0.478  0.0177 -1.88     0.00848  0.00533  0.0998  0.000677        146
> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>,
> #   abs_error_x3 <dbl>, abs_error_x4 <dbl>, abs_error_x5 <dbl>, ...
\end{verbatim}

>>>>

\begin{verbatim}
pak::pak("JayaniLakshika/quollr")
\end{verbatim}

@@

\begin{verbatim}
> $scaled_nldr
> # A tibble: 1,000 x 3
>     emb1  emb2    ID
>    <dbl> <dbl> <int>
>  1 0.277 0.913     1
>  2 0.697 0.538     2
>  3 0.779 0.399     3
>  4 0.173 0.953     4
>  5 0.218 0.983     5
>  6 0.593 1.05      6
>  7 0.180 0.210     7
>  8 0.976 0.571     8
>  9 0.803 0.829     9
> 10 0.932 0.410    10
> # i 990 more rows
> 
> $lim1
> [1] -9.146398  8.552211
> 
> $lim2
> [1] -10.36686  10.10691
\end{verbatim}

@@

\begin{verbatim}
> $b2
> [1] 20
> 
> $a1
> [1] 0.0834108
> 
> $a2
> [1] 0.07223587
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 300 x 3
>        h     c_x    c_y
>    <int>   <dbl>  <dbl>
>  1     1 -0.1    -0.116
>  2     2 -0.0166 -0.116
>  3     3  0.0668 -0.116
>  4     4  0.150  -0.116
>  5     5  0.234  -0.116
>  6     6  0.317  -0.116
>  7     7  0.400  -0.116
>  8     8  0.484  -0.116
>  9     9  0.567  -0.116
> 10    10  0.651  -0.116
> # i 290 more rows
\end{verbatim}

@@

For each centroid, six vertices are computed and assigned a polygon ID (\texttt{h}) corresponding to the centroid's \texttt{h}. These points are aggregated into a tibble that includes the polygon ID (\texttt{h}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.

@@

\begin{verbatim}
>   h          x           y
> 1 1 -0.1000000 -0.06752289
> 2 1 -0.1417054 -0.09160152
> 3 1 -0.1417054 -0.13975877
> 4 1 -0.1000000 -0.16383739
> 5 1 -0.0582946 -0.13975877
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 4
>    emb1  emb2    ID     h
>   <dbl> <dbl> <int> <int>
> 1 0.277 0.913     1   216
> 2 0.697 0.538     2   145
> 3 0.779 0.399     3   116
> 4 0.173 0.953     4   229
> 5 0.218 0.983     5   229
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 3
>       h   n_h   w_h
>   <int> <int> <dbl>
> 1    34     1 0.001
> 2    35    10 0.01 
> 3    36     8 0.008
> 4    37     8 0.008
> 5    38     2 0.002
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 2
>       h pts_list  
>   <int> <list>    
> 1    34 <int [1]> 
> 2    35 <int [10]>
> 3    36 <int [8]> 
> 4    37 <int [8]> 
> 5    38 <int [2]>
\end{verbatim}

@@

\begin{verbatim}
>   h        c_x        c_y n_h w_h
> 1 1 -0.1000000 -0.1156801   0   0
> 2 2 -0.0165892 -0.1156801   0   0
> 3 3  0.0668216 -0.1156801   0   0
> 4 4  0.1502324 -0.1156801   0   0
> 5 5  0.2336432 -0.1156801   0   0
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 8
>    from    to  x_from  y_from    x_to    y_to from_count to_count
>   <int> <int>   <dbl>   <dbl>   <dbl>   <dbl>      <dbl>    <dbl>
> 1     1     2 -0.1    -0.116  -0.0166 -0.116           0        0
> 2    16    17 -0.0583 -0.0434  0.0251 -0.0434          0        0
> 3    16    32 -0.0583 -0.0434 -0.0166  0.0288          0        0
> 4     3    17  0.0668 -0.116   0.0251 -0.0434          0        0
> 5    32    33 -0.0166  0.0288  0.0668  0.0288          0        0
\end{verbatim}

@@

\begin{verbatim}
find_low_dens_hex(
  model_2d = df_bin_centroids, 
  b1 = 15, 
  benchmark_mean_dens = 0.05
)
\end{verbatim}

\begin{verbatim}
>   [1]   1   2   3   4   5   6   7   8   9  10  11  12  13  14  15  16  17  18
>  [19]  19  20  21  22  23  24  25  26  27  28  29  30  31  32  33  34  35  36
>  [37]  37  38  39  40  41  42  43  44  45  46  47  48  49  50  51  52  53  54
>  [55]  55  56  57  58  59  60  61  62  63  64  65  66  67  68  69  70  71  72
>  [73]  73  74  75  76  77  78  79  80  81  82  83  84  85  86  87  88  89  90
>  [91]  91  92  93  94  95  96  97  98  99 100 101 102 103 104 105 106 107 108
> [109] 109 110 111 112 113 114 115 116 117 118 119 120 121 122 123 124 125 126
> [127] 127 128 129 130 131 132 133 134 135 136 137 138 139 140 141 142 143 144
> [145] 145 146 147 148 149 150 151 152 153 154 155 156 157 158 159 160 161 162
> [163] 163 164 165 166 167 168 169 170 171 172 173 174 175 176 177 178 179 180
> [181] 181 182 183 184 185 186 187 188 189 190 191 192 193 194 195 196 197 198
> [199] 199 200 201 202 203 204 205 206 207 208 209 210 211 212 213 214 215 216
> [217] 217 218 219 220 221 222 223 224 225 226 227 228 229 230 231 232 233 234
> [235] 235 236 237 238 239 240 241 242 243 244 245 246 247 248 249 250 251 252
> [253] 253 254 255 256 257 258 259 260 261 262 263 264 265 266 267 268 269 270
> [271] 271 272 273 274 275 276 277 278 279 280 281 282 283 284 285 286 287 288
> [289] 289 290 291 292 293 294 295 296 297 298 299 300
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 8
>       h     x1     x2    x3         x4        x5      x6         x7
>   <int>  <dbl>  <dbl> <dbl>      <dbl>     <dbl>   <dbl>      <dbl>
> 1    34 0.958  0.0854  1.29  0.00265    0.0171    0.0876 -0.00249  
> 2    35 0.818  0.116   1.56  0.00184    0.00361  -0.0318 -0.00377  
> 3    36 0.544  0.111   1.83 -0.00341   -0.000303  0.0196  0.0000704
> 4    37 0.279  0.128   1.95  0.0000880  0.00104  -0.0276 -0.000227 
> 5    38 0.0567 0.119   2.00  0.00733    0.00238   0.0833 -0.00166
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 4
>   pred_emb_1 pred_emb_2    ID pred_h
>        <dbl>      <dbl> <int>  <int>
> 1         NA         NA     1    215
> 2         NA         NA     2    145
> 3         NA         NA     3    116
> 4         NA         NA     4    229
> 5         NA         NA     5    229
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 1 x 2
>   Error  RMSE
>   <dbl> <dbl>
> 1  265. 0.157
\end{verbatim}

@@

\begin{verbatim}
> # A tibble: 5 x 32
>      ID      x1    x2       x3       x4       x5      x6       x7 pred_h
>   <int>   <dbl> <dbl>    <dbl>    <dbl>    <dbl>   <dbl>    <dbl>  <int>
> 1     1 -0.120  0.819 -1.99     0.0114   0.00351  0.0334  0.00638    215
> 2     2 -0.0492 0.166  0.00121  0.0115  -0.0166  -0.0297  0.00509    145
> 3     3 -0.774  0.651  0.367   -0.0172   0.00600  0.0211  0.00303    116
> 4     4 -0.606  0.952 -1.80     0.0157  -0.00978 -0.0590 -0.00754    229
> 5     5 -0.478  1.10  -1.88    -0.00423  0.00495 -0.0482 -0.00982    229
> # i 23 more variables: model_high_d_x1 <dbl>, model_high_d_x2 <dbl>,
> #   model_high_d_x3 <dbl>, model_high_d_x4 <dbl>, model_high_d_x5 <dbl>,
> #   model_high_d_x6 <dbl>, model_high_d_x7 <dbl>, error_square_x1 <dbl>,
> #   error_square_x2 <dbl>, error_square_x3 <dbl>, error_square_x4 <dbl>,
> #   error_square_x5 <dbl>, error_square_x6 <dbl>, error_square_x7 <dbl>,
> #   row_wise_total_error <dbl>, abs_error_x1 <dbl>, abs_error_x2 <dbl>,
> #   abs_error_x3 <dbl>, abs_error_x4 <dbl>, abs_error_x5 <dbl>, ...
\end{verbatim}
